---
title: "flat_full.Rmd for working package"
output: html_document
editor_options: 
  chunk_output_type: console
---

<!-- Run this 'development' chunk -->
<!-- Store every call to library() that you need to explore your functions -->

```{r development, include=FALSE}

#Paquetes necesarios
pacman::p_load(
    #test
    pkgload, pkgdown, testthat,
    #data tools
    tidyverse, janitor, glue, lubridate, scales, Hmisc,lazyeval,
    #Gráficos
    plotly, ggrepel, cowplot, grid, fmsb, 
    #importar / exportar
    haven, rio, officer, officedown,
    #Etiquetas
    sjlabelled,
    #tablas
    flextable, knitr, kableExtra, DT, gtsummary,
    #Temas
    ggpubr,
    #Colores
    paletteer, RColorBrewer
    )

library(pulso2)

```

<!--
 You need to run the 'description' chunk in the '0-dev_history.Rmd' file before continuing your code there.

If it is the first time you use {fusen}, after 'description', you can directly run the last chunk of the present file with inflate() inside.
--> 

<!-- 
 Store your dataset in a directory named "inst/" at the root of your project.
 Use it for your tests in this Rmd thanks to `pkgload::load_all()` to make it available
and `system.file()` to read it in your examples.

- There already is a dataset in the "inst/" directory to be used in the examples below
-->

```{r development-2}
# Run all this chunk in the console directly
# There already is a dataset in the "inst/" directory
# Make the dataset file available to the current Rmd during development
pkgload::load_all(path = here::here(), export_all = FALSE)

# You will be able to read your example data file in each of your function examples and tests as follows - see chunks below
#dir.create(here::here("inst"))

# data_prueba_ruta <- system.file("data_prueba.sav", package = "pulso2")
# data_prueba <- haven::read_sav(data_prueba_ruta)

```

# Empezando...

Para estos ejemplos se utilizara una base aleatoria de un estudio realizado por Pulso PUCP el cual nos permitira mostrar la manera en que se puede utilizar el paquete `pulso`

# Barra simple

```{r function-barra_simple}
#' Funcion para realizar un grafico de barras simple.
#' @import testthat
#' @import qpdf
#' @import pkgload
#' @import pkgdown
#' 
#' @importFrom haven read_sav
#' @importFrom dplyr filter count mutate select summarise arrange
#' @importFrom forcats fct_rev fct_infreq fct_relevel
#' @importFrom sjlabelled as_label
#' @importFrom scales number_format label_wrap wrap_format number percent
#' @importFrom ggplot2 ggplot geom_bar geom_text scale_x_discrete scale_y_continuous coord_flip labs theme aes position_dodge element_text element_line element_blank position_stack waiver
#' @importFrom janitor round_half_up
#' @importFrom ggpubr theme_pubr
#' @importFrom glue glue
#'
#' @param data Base de datos para la funcion
#' @param var Variable para el grafico de barras simple
#' @param filtrar Por default TRUE que significa ningún filtro, si se quiere filtrar por una variable especificar la variable y el valor a filtrar (ej: filtrar = q0002 == 1 que significa filtrar la base para que aparezcan solo los casos que tienen en la pregunta q0002 el valor de 1).
#' @param ordenado Por default es TRUE, significa que ordena las barras de menor a mayor, cambiar a FALSE para dejarlas en el orden original.
#' @param frecuencia Por default es FALSE, cambiar a TRUE para visualizar los resultados en frecuencias y no en porcentajes.
#' @param porcentaje Por default es TRUE, cambiar a FALSE para visualizar los resultados de porcentajes sin el simbolo %.
#' @param ultimo Por default es NULL, escribir entre comillas la palabra que quieres que vaya al final de las barras (ej: ultimo="Otros")
#' @param max.limit Por default es 1, es el limite máximo del eje x. 1=100 cuando son resultados en porcentajes (ej: max.limit=0.75 significa maximo 75%). Aunque el porcentaje esté en FALSE, el max.limit debe ser tratado como si 1.00 fuera el maximo y no como si 100 fuera el maximo.
#' @param color Por default es color azul que es "#B0D597" en notacion hexagesimal. Buscar colores hex para más información. También permite colores grabados en R como "red".
#' @param ext.label Por default es 30, a menor el número menor espacio para el texto de las etiquetas.
#'
#' @return
#' Un grafico de barras simple.
#' 
#' @examples
#' @export

barra_simple<- function(data, var, filtrar=TRUE, ordenado=TRUE, frecuencia=FALSE,  porcentaje=TRUE, ultimo=NULL, max.limit=1, color = "#B0D597", ext.label=30){
  
  total<-nrow(data)
  
  if(isTRUE(ordenado)) {
    
    tablon<-data %>%
      filter({{filtrar}}) %>% 
      as_label() %>%
      count(var = fct_rev(fct_infreq(factor({{var}}))) ) %>%
      mutate(pct = prop.table(n))
    
  } else {
    
    tablon<-data %>%
      filter({{filtrar}}) %>% 
      as_label() %>%
      count(var = fct_rev(factor({{var}}))) %>%
      mutate(pct = prop.table(n))
    
  }
  
  if(isTRUE(frecuencia)) {
    
    tablon %>%
      #grafico
      ggplot(aes(x = fct_relevel(var, ultimo), y = n ) ) +
      geom_bar(stat='identity', fill = color, width = 0.6) +
      
      #Etiqueta = -7%
      geom_text(aes(label = ifelse(pct < 0.07, n, "") ),
                position = position_dodge(width = .9),
                vjust = 0.2,
                hjust = -0.2,
                size = 3.5,
                fontface = "bold",
                color = "#002060") +
      
      #Etiqueta = El resto
      geom_text(aes(label = ifelse(pct >= 0.07, n, "") ),
                position = position_stack(vjust = 0.5),
                size = 3.5,
                fontface = "bold",
                color = "#002060") +
      
      scale_x_discrete(labels = wrap_format(ext.label)) +
      scale_y_continuous(labels=number_format(accuracy =1), limits = c(0, if(max.limit==1){max(tablon$n) + round_half_up(max(tablon$n)/3)} else {max.limit} )) +
      coord_flip() +
      theme_pubr() +
      labs(subtitle = "Resultados en frecuencias",
           caption = "Elaborado por Pulso PUCP",
           tag = if(sum(tablon$n) == total){glue("N=",sum(tablon$n))}else{glue("N=",sum(tablon$n),"/",total)}) +
      theme(text = element_text(size = 9, color="#002060"),
            
            plot.subtitle = element_text(size = 10, color="#002060"),
            plot.title.position = "plot",
            
            plot.caption = element_text(face = "italic"),
            
            plot.tag = element_text(size = 8, color="grey40"),
            plot.tag.position = "topright",
            
            axis.title = element_blank(),
            axis.text = element_text(color="#002060"),
            axis.ticks = element_line(color="#002060"),
            axis.line = element_line(color="#002060", size = 0.5) )
    
  } else {
    
    tablon %>%
      #grafico
      ggplot(aes(x = fct_relevel(var, ultimo), y = pct ) ) +
      geom_bar(stat='identity', fill = color, width = 0.6) +
      
      #Etiqueta = -7%
      geom_text(aes(label = ifelse(pct < 0.07, if(isTRUE(porcentaje)){scales::percent(pct, accuracy = 1)} else {scales::number(pct, scale = 100)} , "") ),
                position = position_dodge(width = .9),
                vjust = 0.2,
                hjust = -0.2,
                size = 3.5,
                fontface = "bold",
                color = "#002060") +
      
      #Etiqueta = El resto
      geom_text(aes(label = ifelse(pct >= 0.07, if(isTRUE(porcentaje)){scales::percent(pct, accuracy = 1)} else {scales::number(pct, scale = 100)}, "") ),
                position = position_stack(vjust = 0.5),
                size = 3.5,
                fontface = "bold",
                color = "#002060") +
      
      scale_x_discrete(labels = wrap_format(ext.label)) +
      scale_y_continuous(labels = if(isTRUE(porcentaje)) {~scales::percent(.x, accuracy = 1)} else {~scales::number(.x, scale = 100)}, limits = if(isTRUE(porcentaje)) {c(0, max.limit)} else { c(0, max.limit )} ) +
      coord_flip() +
      theme_pubr() +
      labs(subtitle = if(isTRUE(porcentaje)){waiver()} else {"Resultados en porcentajes"},
           caption = "Elaborado por Pulso PUCP",
           tag = if(sum(tablon$n) == total){glue("N=",sum(tablon$n))}else{glue("N=",sum(tablon$n),"/",total)}) +
      theme(text = element_text(size = 9, color="#002060"),
            
            plot.subtitle = element_text(size = 10, color="#002060"),
            plot.title.position = "plot",
            
            plot.caption = element_text(face = "italic"),
            
            plot.tag = element_text(size = 8, color="grey40"),
            plot.tag.position = "topright",
            
            axis.title = element_blank(),
            axis.text = element_text(color="#002060"),
            axis.ticks = element_line(color="#002060"),
            axis.line = element_line(color="#002060", size = 0.5) )
      
  }
  
}

```

```{r examples-barra_simple}
#library(pulso2)

#importar base de datos
data_prueba_ruta <- system.file("data_prueba.sav", package = "pulso2")
data_prueba <- haven::read_sav(data_prueba_ruta)

#aplicar funcion
data_prueba |>
  barra_simple(gedad, ordenado=FALSE)

```

```{r tests-barra_simple}

# #importar base de datos
# data_prueba_ruta <- system.file("data_prueba.sav", package = "pulso")
# data_prueba <- haven::read_sav(data_prueba_ruta)
# 
# test_that("barra_simple works", {expect_error(data_prueba %>% barra_simple(gedad, ordenado=FALSE), regexp = NA)})


```

<!--
Create a chunk for the core of the function

- The chunk needs to be named `function` at least
- It contains the code of a documented function
- The chunk can also be named `function-my_median` to make it easily
findable in your Rmd
- Let the `@examples` part empty, and use the next `examples` chunk instead to present reproducible examples

After inflating the template

-  This function code will automatically be added in a new file in the "R/" directory
-->

```{r function}
#' My median
#' @import testthat
#' @import qpdf
#'
#' @param x Vector of Numeric values
#' @inheritParams stats::median
#'
#' @return
#' Median of vector x
#' @export
#'
#' @examples
my_median <- function(x, na.rm = TRUE) {
  if (!is.numeric(x)) {stop("x should be numeric")}
  stats::median(x, na.rm = na.rm)
}
```

<!--
Create a chunk with an example of use for your function

- The chunk needs to be named `examples` at least
- It contains working examples of your function
- The chunk is better be named `examples-my_median` to be handled
correctly when inflated as a vignette

After inflating the template

-  This example will automatically be added in the '@examples' part of our function above in the "R/" directory
- This example will automatically be added in the vignette created from this Rmd template
-->

```{r examples}
my_median(1:12)

# Example with your dataset in "inst/"
datafile <- system.file("nyc_squirrels_sample.csv", package = "pulso2")
nyc_squirrels <- read.csv(datafile)
# Apply my function
my_median(nyc_squirrels[,"hectare_squirrel_number"])
```

<!--
Create a chunk with a test of use for your function

- The chunk needs to be named `tests` at least
- It contains working tests of your function
- The chunk is better be named `tests-my_median` to be handled
correctly when inflated as a vignette

After inflating the template

-  This test code will automatically be added in the "tests/testthat/" directory
-->

```{r tests}
test_that("my_median works properly and show error if needed", {
  expect_true(my_median(1:12) == 6.5)
  expect_error(my_median("text"))
})

# Test with your dataset in "inst/"
datafile <- system.file("nyc_squirrels_sample.csv", package = "pulso2")
nyc_squirrels <- read.csv(datafile)
# Apply test on my function
test_that("my_median works properly with internal dataset", {
  expect_equal(my_median(nyc_squirrels[,"hectare_squirrel_number"]), 3)
})
```

# Calculate the mean of a vector

<!--
There can be other functions, examples and tests in your flat template.
Each of them will be inflated in a different file, provided that there is a level-1 or level-2 section title to separate from previous functions.
-->

## Use sub-functions in the same chunk

```{r function-my_other_median}
#' My Other median
#'
#' @param x Vector of Numeric values
#' @inheritParams stats::median
#'
#' @return
#' Median of vector x
#' @export
#'
#' @examples
my_other_median <- function(x, na.rm = TRUE) {
  if (!is.numeric(x)) {stop("x should be numeric")}
  sub_median(x, na.rm =na.rm)
}

#' Core of the median not exported
#' @param x Vector of Numeric values
#' @inheritParams stats::median
sub_median <- function(x, na.rm = TRUE) {
  stats::median(x, na.rm)
}
```

```{r examples-my_other_median}
my_other_median(1:12)
```

```{r tests-my_other_median}
test_that("my_median works properly and show error if needed", {
  expect_true(my_other_median(1:12) == 6.5)
  expect_error(my_other_median("text"))
})
```

<!--
# There can be development actions

Create a chunk with 'development' actions

- The chunk needs to be named `development` or `dev`
- It contains functions that are used for package development only
- Note that you may want to store most of these functions in the 0-dev_history.Rmd file

These are only included in the present flat template file, their content will not be part of the package anywhere else.
-->

```{r development-inflate, eval=FALSE}
# Keep eval=FALSE to avoid infinite loop in case you hit the knit button
# Execute in the console directly
Sys.setlocale(category = "LC_ALL", locale = "Spanish_Spain.1252")
fusen::inflate(flat_file = "dev/flat_full.Rmd", vignette_name = "Empezando")
```


# Inflate your package

You're one inflate from paper to box.
Build your package from this very Rmd using `fusen::inflate()`

- Verify your `"DESCRIPTION"` file has been updated
- Verify your function is in `"R/"` directory
- Verify your test is in `"tests/testthat/"` directory
- Verify this Rmd appears in `"vignettes/"` directory
